---
title: "business_forecasting_assignment05"
author: "Adwait Kalsekar"
date: "2023-10-27"
output: html_document
---

# Dataset Information

The dataset chosen for this assignment consists of the House Price Index for the State of New Jersey over the last 50 years calculated quarterly.

The time series of the data set shows seasonality in the House Price Index every 10 years. The HPI grows exponentially over the decade and drops towards the start of a new decade. This seasonality holds true for the entire time series data.

Looking at the time series data it can also be inferred that there has been a visible trend in the growth of HPI over the last 50 years. The HPI has increased exponentially since the year 1975 to the year 2023. 

Looking at the time series data, certain forecasting models applied have predicted similar future data points that account for the existing trend and seasonality.


## Loading the Dataset and Converting to Time Series

```{r}
# Loading Libraries
library(fpp)
library(fpp2)
library(TTR)

# Reading Data from CSV File
df = read.csv("NJ_HPI.csv")
str(df)

# Converting DATE column to R Date Type
df$DATE <- as.Date(df$DATE)
str(df)
df

# Converting to Time Series data
df_ts = ts(df$HPI.NJ, start=c(1975, 1), end=c(2023, 2), frequency=4)
df_ts

# Attributes and Structure
attributes(df_ts)
str(df_ts)

```

## Plotting and Observing the Time Series

```{r}
summary(df_ts)
plot(df_ts)

```

# Decomposition

## Seasonal Decomposition of Time Series by Loess (STL) Decomposition

```{r}

# STL Decomposition
df_ts_stl <- stl(df_ts,s.window ="periodic")
attributes(df_ts_stl)
plot(df_ts_stl)

```
## Plotting Trend and Seasonal Components
```{r}
plot(df_ts_stl$time.series[, "trend"], main="Trend Component")
plot(df_ts_stl$time.series[, "seasonal"], main="Seasonal Component")

```

## Seasonally Adjusting Time Series

```{r}
# Seasonal Adjustment
df_ts_stl_seasadj = seasadj(df_ts_stl)
plot(df_ts_stl_seasadj)
```

## Plotting Line for Seasonally Adjusted STL with the original Time Series

```{r}
plot(df_ts)
lines(df_ts_stl_seasadj, col="blue")
```

### As the data is not seasonal, there is no difference between seasonally adjusted time series and original time series and the two lines overlap 

## Forecasting the Decomposed Time Series

```{r}
# Default period forecast
df_ts_stl_forecast <- forecast(df_ts_stl)
plot(df_ts_stl_forecast)

# Passing number of periods
df_ts_stl_forecast_20 <- forecast(df_ts_stl, h=20)
plot(df_ts_stl_forecast_20)
```

## Forecasting the Seasonally Adjusted Decomposed Time Series

```{r}
df_ts_stl_seasadj_forecast <- forecast(df_ts_stl_seasadj, h=20)
plot(df_ts_stl_seasadj_forecast)
```

## Using Decompose Function

```{r}
df_ts_decompose <- decompose(df_ts)
attributes(df_ts_decompose)
plot(df_ts_decompose)

# Seasonal Adjustment
df_ts_decompose_seasadj = seasadj(df_ts_decompose)
plot(df_ts_decompose_seasadj)
```

## Plotting Line Seasonally Adjusted Decomposed TS with the original Time Series

```{r}

plot(df_ts)
lines(df_ts_decompose_seasadj, col="blue")
```

## Forecasting the Seasonally Adjusted Decomposed Time Series

```{r}
df_ts_decompose_seasadj_forecast <- forecast(df_ts_decompose_seasadj, h=20)
plot(df_ts_decompose_seasadj_forecast)
```
